---
title: "Shopify Sales Dashboard"
author: "Ron Stephen Sabilla"
format:
  dashboard:
    theme: cosmo
    scrolling: true
    nav-buttons:
      - icon: github
        href: https://github.com
logo: ""
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
library(lubridate)
library(janitor)
library(plotly)
library(DT)
library(bslib)

# ── Load & clean ──────────────────────────────────────────────────────────────
# Place your downloaded CSV (from Kaggle) in the same folder as this .qmd file
# and update the filename below if needed.
df_raw <- read_csv("shopify_sales.csv", show_col_types = FALSE)

# Normalize column names to snake_case so the code below works regardless
# of how Kaggle capitalizes them
df <- df_raw |> clean_names()

# ── Column name mapping ───────────────────────────────────────────────────────
# Update the RIGHT-HAND side values to match your CSV's snake_case column names
col_map <- c(
  order_date = "created_at",
  revenue    = "total_price",
  category   = "product_type",
  product    = "lineitem_name",
  quantity   = "lineitem_quantity",
  status     = "fulfillment_status",
  payment    = "payment_method",
  country    = "billing_country",
  city       = "billing_city"
)

# ── Rename to standard names used throughout ──────────────────────────────────
# any_of() safely skips columns that don't exist in this dataset
df <- df |>
  rename(any_of(col_map)) |>
  mutate(
    order_date = parse_date_time(order_date,
                   orders = c("ymd HMS", "ymd HM", "ymd", "mdy", "dmy"),
                   quiet  = TRUE),
    order_day  = as_date(order_date),
    order_hour = hour(order_date),
    order_wday = wday(order_date, label = TRUE, abbr = TRUE),
    across(any_of(c("revenue", "quantity")), as.numeric)
  ) |>
  filter(!is.na(revenue), revenue >= 0)

# ── KPI values ────────────────────────────────────────────────────────────────
total_revenue   <- sum(df$revenue, na.rm = TRUE)
total_orders    <- n_distinct(df$order_id %||% seq_len(nrow(df)))
avg_order_value <- total_revenue / max(total_orders, 1)
total_units     <- sum(df$quantity, na.rm = TRUE)

# Colour palette
clr <- c("#5C4DE5", "#00B4D8", "#0077B6", "#48CAE4",
         "#90E0EF", "#ADE8F4", "#CAF0F8", "#03045E")
```

# Overview

## Row {height=20%}

```{r}
#| content: valuebox
#| title: "Total Revenue"
list(
  icon  = "currency-dollar",
  color = "primary",
  value = dollar(total_revenue, accuracy = 1)
)
```

```{r}
#| content: valuebox
#| title: "Total Orders"
list(
  icon  = "bag-check",
  color = "info",
  value = comma(total_orders)
)
```

```{r}
#| content: valuebox
#| title: "Avg Order Value"
list(
  icon  = "graph-up-arrow",
  color = "success",
  value = dollar(avg_order_value, accuracy = 0.01)
)
```

```{r}
#| content: valuebox
#| title: "Units Sold"
list(
  icon  = "box-seam",
  color = "warning",
  value = comma(total_units)
)
```

## Row {height=40%}

### Revenue Over Time {width=60%}

```{r}
daily_rev <- df |>
  group_by(order_day) |>
  summarise(revenue = sum(revenue, na.rm = TRUE), .groups = "drop")

p1 <- ggplot(daily_rev, aes(x = order_day, y = revenue)) +
  geom_area(fill = "#5C4DE5", alpha = 0.2) +
  geom_line(color = "#5C4DE5", linewidth = 1) +
  scale_y_continuous(labels = dollar_format(scale = 1e-3, suffix = "K")) +
  scale_x_date(date_labels = "%b %d") +
  labs(x = NULL, y = "Daily Revenue") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank())

ggplotly(p1, tooltip = c("x", "y")) |>
  layout(hovermode = "x unified")
```

### Orders by Status {width=40%}

```{r}
if ("status" %in% names(df)) {
  status_tbl <- df |>
    count(status) |>
    filter(!is.na(status)) |>
    mutate(status = str_to_title(status))

  plot_ly(status_tbl,
    labels = ~status, values = ~n,
    type   = "pie",
    hole   = 0.5,
    marker = list(colors = clr),
    textinfo = "label+percent"
  ) |>
    layout(
      showlegend = TRUE,
      legend = list(orientation = "h")
    )
} else {
  cat("No 'fulfillment_status' column found — update status_col in setup.")
}
```

## Row {height=40%}

### Sales by Product Category {width=50%}

```{r}
if ("category" %in% names(df)) {
  cat_rev <- df |>
    group_by(category) |>
    summarise(revenue = sum(revenue, na.rm = TRUE), .groups = "drop") |>
    filter(!is.na(category)) |>
    arrange(desc(revenue)) |>
    slice_head(n = 10) |>
    mutate(category = fct_reorder(category, revenue))

  p2 <- ggplot(cat_rev, aes(x = category, y = revenue, fill = category)) +
    geom_col(show.legend = FALSE) +
    scale_fill_manual(values = colorRampPalette(clr)(nrow(cat_rev))) +
    scale_y_continuous(labels = dollar_format(scale = 1e-3, suffix = "K")) +
    coord_flip() +
    labs(x = NULL, y = "Revenue") +
    theme_minimal(base_size = 13) +
    theme(panel.grid.minor = element_blank())

  ggplotly(p2, tooltip = c("x", "y"))
} else {
  cat("No category column found — update category_col in setup.")
}
```

### Peak Sales by Hour {width=50%}

```{r}
hourly <- df |>
  group_by(order_hour) |>
  summarise(
    orders  = n(),
    revenue = sum(revenue, na.rm = TRUE),
    .groups = "drop"
  )

p3 <- ggplot(hourly, aes(x = order_hour, y = orders)) +
  geom_col(fill = "#00B4D8", alpha = 0.85) +
  scale_x_continuous(breaks = 0:23, labels = paste0(0:23, "h")) +
  labs(x = "Hour of Day", y = "Number of Orders") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggplotly(p3, tooltip = c("x", "y"))
```

# Products

## Row {height=50%}

### Top 10 Products by Revenue

```{r}
if ("product" %in% names(df)) {
  top_products <- df |>
    group_by(product) |>
    summarise(
      revenue = sum(revenue, na.rm = TRUE),
      orders  = n(),
      units   = sum(quantity, na.rm = TRUE),
      .groups = "drop"
    ) |>
    filter(!is.na(product)) |>
    arrange(desc(revenue)) |>
    slice_head(n = 10) |>
    mutate(product = fct_reorder(product, revenue))

  p4 <- ggplot(top_products, aes(x = product, y = revenue, fill = revenue)) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(low = "#90E0EF", high = "#03045E") +
    scale_y_continuous(labels = dollar_format(scale = 1e-3, suffix = "K")) +
    coord_flip() +
    labs(x = NULL, y = "Revenue") +
    theme_minimal(base_size = 13)

  ggplotly(p4, tooltip = c("x", "y"))
} else {
  cat("No product column found — update product_col in setup.")
}
```

### Sales by Day of Week {width=40%}

```{r}
wday_rev <- df |>
  group_by(order_wday) |>
  summarise(
    revenue = sum(revenue, na.rm = TRUE),
    orders  = n(),
    .groups = "drop"
  )

p5 <- ggplot(wday_rev, aes(x = order_wday, y = revenue, fill = order_wday)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = colorRampPalette(clr)(7)) +
  scale_y_continuous(labels = dollar_format(scale = 1e-3, suffix = "K")) +
  labs(x = NULL, y = "Revenue") +
  theme_minimal(base_size = 13)

ggplotly(p5, tooltip = c("x", "y"))
```

## Row {height=50%}

### Full Product Table

```{r}
if ("product" %in% names(df)) {
  df |>
    group_by(Product = product) |>
    summarise(
      Revenue  = dollar(sum(revenue, na.rm = TRUE)),
      Orders   = comma(n()),
      Units    = comma(sum(quantity, na.rm = TRUE)),
      Avg_Price = dollar(mean(revenue, na.rm = TRUE)),
      .groups = "drop"
    ) |>
    arrange(desc(parse_number(Revenue))) |>
    datatable(
      options = list(pageLength = 10, scrollX = TRUE),
      rownames = FALSE,
      class = "compact stripe hover"
    )
}
```

# Geography

## Row

### Revenue by Country

```{r}
if ("country" %in% names(df)) {
  country_rev <- df |>
    group_by(country) |>
    summarise(revenue = sum(revenue, na.rm = TRUE), .groups = "drop") |>
    filter(!is.na(country)) |>
    arrange(desc(revenue))

  plot_ly(country_rev,
    type  = "choropleth",
    locations = ~country,
    locationmode = "country names",
    z     = ~revenue,
    colorscale = "Blues",
    colorbar = list(title = "Revenue ($)")
  ) |>
    layout(
      geo = list(showframe = FALSE, showcoastlines = TRUE)
    )
} else {
  cat("No country column found — update country_col in setup.")
}
```

## Row

### Top 10 Cities by Revenue {width=50%}

```{r}
if ("city" %in% names(df)) {
  city_rev <- df |>
    group_by(city) |>
    summarise(revenue = sum(revenue, na.rm = TRUE), .groups = "drop") |>
    filter(!is.na(city)) |>
    arrange(desc(revenue)) |>
    slice_head(n = 10) |>
    mutate(city = fct_reorder(city, revenue))

  p6 <- ggplot(city_rev, aes(x = city, y = revenue, fill = revenue)) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(low = "#90E0EF", high = "#03045E") +
    scale_y_continuous(labels = dollar_format(scale = 1e-3, suffix = "K")) +
    coord_flip() +
    labs(x = NULL, y = "Revenue") +
    theme_minimal(base_size = 13)

  ggplotly(p6, tooltip = c("x", "y"))
} else {
  cat("No city column found — update city_col in setup.")
}
```

### Payment Methods {width=50%}

```{r}
if ("payment" %in% names(df)) {
  pay_tbl <- df |>
    count(payment) |>
    filter(!is.na(payment)) |>
    mutate(payment = str_to_title(payment))

  plot_ly(pay_tbl,
    labels = ~payment, values = ~n,
    type   = "pie",
    hole   = 0.4,
    marker = list(colors = clr),
    textinfo = "label+percent"
  ) |>
    layout(showlegend = TRUE)
} else {
  cat("No payment method column found — update payment_col in setup.")
}
```

# About

## Row

::: {.card title="About This Dashboard"}

**Dataset:** Shopify Sales Dataset for ML & EDA
**Source:** [Kaggle — aliiihussain](https://www.kaggle.com/datasets/aliiihussain/shopify-sales-dataset-for-ml-and-eda)
**Description:** A realistic e-commerce dataset containing Shopify order-level data for sales analysis and customer behavior insights.

### Story

This dashboard tells the story of **where revenue comes from, when customers buy, and which products drive the most value** for a Shopify-based e-commerce store. Even someone unfamiliar with e-commerce can use these four pages to answer the most important business questions:

- **Overview** — How is the store performing overall? What are the peak hours and order fulfillment health?
- **Products** — Which products and categories generate the most revenue? Are there slow days that need promotion?
- **Geography** — Where in the world are customers buying from? Which cities are highest-value?
- **About** — Context and data source.

### Key Takeaways

1. A small number of product categories typically drive the majority of revenue (Pareto effect).
2. Order volume peaks at specific hours of the day, suggesting optimal times for email campaigns.
3. Certain days of the week consistently outperform others — useful for scheduling promotions.
4. Geographic concentration reveals where marketing spend is most effective.

### How to Use

Download the CSV from Kaggle, place it in the same folder as this `.qmd` file, and render with:

```bash
quarto render shopify_dashboard.qmd
```

If column names in your CSV differ from the defaults, update the mapping section at the top of the setup chunk.
:::
